---
title: "Application exercise - Insights from Crime Data in NZ"
author: "Andres Ponce"
output:
 html_notebook:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: true  ## if you want number sections at each table header
    theme: united  # many options for theme, this one is my favorite.
    fig_caption: yes
---

```{r include=FALSE}

#install.packages("readxl")
library(readxl)
#install.packages("WDI")
library(tidyverse)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("plotly")
library(plotly)
#install.packages("DT")
library(DT)
library(gridExtra)
library(grid)
```


```{r include=FALSE}
#Preparting the Data
#crime <- read_excel("D:/Behavioral NZ/BI_recruitment_exercise_data.xls")
crime <- read_excel("/media/andres/E859-0BC8/Behavioral NZ/BI_recruitment_exercise_data.xls")
crime <- crime[-c(1, 2, 21), ]


#creating separate datasets for regions and Auckland
crime_total_regions<-dplyr::select(crime, 1:36)
crime_total_Auckland<-dplyr::select(crime,1, 37:71)

#using the first row as variable names
names(crime_total_regions) <- lapply(crime_total_regions[1, ], as.character)
crime_total_regions <- crime_total_regions[-c(1), ]

names(crime_total_Auckland)<- lapply(crime_total_Auckland[1,], as.character)
crime_total_Auckland<- crime_total_Auckland[-c(1),]

#gather data from years as values 
crime_total_Auckland<- crime_total_Auckland %>% gather(year, convictions, 2:36) 
crime_total_regions<- crime_total_regions %>% gather(year, convictions, 2:36)

#parse variables convictions to numeric
crime_total_Auckland$convictions<- crime_total_Auckland$convictions %>% as.numeric()
crime_total_regions$convictions<- crime_total_regions$convictions %>% as.numeric()

#arrange divisions per year
Auckland<-crime_total_Auckland %>% arrange(`ANZSOC division`)
regions<- crime_total_regions %>% arrange(`ANZSOC division`)

#aggregating variables
aggAuckland <-aggregate(Auckland$convictions, by=list(Auckland$`ANZSOC division`), 
  FUN=mean, na.rm=TRUE)

aggRegions <- aggregate(regions$convictions, by=list(regions$`ANZSOC division`),
                        FUN=mean, na.rm=TRUE)

```

```{r include=FALSE}
# aggregate data barplot for regions and Auckland

aggAuckland<-aggAuckland%>%
arrange(desc(x))

p<- ggplot(data=aggAuckland, aes(x=Group.1, y=x, fill = Group.1, label=sprintf("%0.2f", round(x, digits = 1)))) +
geom_bar(stat="identity", show.legend = F)+
  labs(x= ' Type of conviction', y= NULL) +
  ggtitle(paste('Auckland'))+
  geom_text(size = 4.5, hjust= -.1,  position = position_dodge(width= 0.9))+
scale_colour_gradient2()+
coord_flip()+
scale_x_discrete(limits = aggAuckland$Group.1)+
theme_classic()+ theme(
  axis.title.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 16))

```

```{r include=FALSE}

# average analysis for regions
aggRegions<-aggRegions%>%
arrange(desc(x))

r<- ggplot(data=aggRegions, aes(x=Group.1, y=x, fill = Group.1, label=sprintf("%0.2f", round(x, digits = 1)))) +
geom_bar(stat="identity", show.legend =F)+
  labs(y = "Average number of convictions", x= ' Type of conviction') +
  ggtitle(paste('Regions'))+
  geom_text(size = 4.5, hjust= -.1,  position = position_dodge(width= 0.9))+
scale_colour_gradient2()+
coord_flip()+
scale_x_discrete(limits = aggRegions$Group.1)+
theme_classic()+ theme(
  axis.title.x = element_text(size = 16),
  axis.text.y = element_text(size = 16),
  axis.title.y = element_text(size = 16))
#for eliminating value labels
#theme(axis.text.y=element_blank())

```

# Analysis of Average Data Over Time

The data used in this analysis depicts the situation of crime convictions of Adults in Auckland and Regions of New zealand, the period correspond to 1980 until 2014. 

The aim of the following graph is to provide an aggregate perspective on different types of offenses and showcase the main differences among the locations.


```{r echo=FALSE, fig.height=20, fig.width=25}
gridExtra::grid.arrange(p, r, nrow=2, top = textGrob("Average number of yearly convictions in Auckland and Regions (1980-2014)", gp= gpar(fontsize=25, font=3)))

```

# Insights On Over-Time Conviction Development

Another way to visualize trends in the conviction data, is to plot the average growth rate of each type of offense. Intuitively, the following plot showcase whether, on average, a specific offense has decreased or increased over time.

The main policy insight is the case of criminal convictions for both locations regarding drug crimes, which could be closely related with the increase of other types of offenses, such as **Public Order**, **Theft**, and **Acts Intented to Cause Injury**.  


```{r include=FALSE}
#Diverging plot for growth rates

#calculating conviction growth rates for Auckland
growth_rate <- function(x)(((x-lag(x))/x))*100

Auckland$growth_rate <- ifelse(Auckland$year==1980,0, growth_rate(Auckland$convictions)) 

#Calculating conviction growth rates for Regions
regions$growth_rate <- ifelse(regions$year== 1980,0, growth_rate(regions$convictions)) 

#aggregating variable Growth Rate of Auckland
GR_Auckland <-aggregate(Auckland$growth_rate, by=list(Auckland$`ANZSOC division`), 
  FUN=mean, na.rm=TRUE)
#erasing total row
GR_Auckland <- GR_Auckland[-c(15), ]

#aggregating Grwth Rate of Regions
GR_Regions <- aggregate(regions$growth_rate, by=list(regions$`ANZSOC division`),
                        FUN=mean, na.rm=TRUE)
# Erasing total row of Regions
GR_Regions <- GR_Regions[-c(15), ]

#Preparing data for ploting Auckland
GR_Auckland$mpg_type <- ifelse(GR_Auckland$x < 0, "below", "above")  # above / below avg flag
GR_Auckland<-GR_Auckland %>% arrange(x)
GR_Auckland$`car name` <- factor(GR_Auckland$Group.1, levels = GR_Auckland$Group.1)  # convert to factor to retain sorted order in plot.

# Plot
GR_A <- ggplot(GR_Auckland, aes(x=`car name`, y=x, label=sprintf("%0.2f", round(x, digits = 1)))) + 
  geom_point(stat='identity', aes(col=mpg_type), size=10)  +
  labs(y = "Average Growth Rate", x= ' Type of conviction')+
  scale_color_manual(name="Average Growth Rate of Convictions", 
                     labels = c("Increase", "Decrease"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="black", size=4) +
  labs(title="Auckland") + 
  ylim(-14, 2) +
  coord_flip()+ theme(axis.text=element_text(size=11))

#Preparing data for ploting Regions
GR_Regions$mpg_type <- ifelse(GR_Regions$x < 0, "below", "above")  # above / below avg flag
GR_Regions<-GR_Regions %>% arrange(x)
GR_Regions$`car name` <- factor(GR_Regions$Group.1, levels = GR_Regions$Group.1)  # convert to factor to retain sorted order in plot.

# Plot
GR_R <- ggplot(GR_Regions, aes(x=`car name`, y=x, label=sprintf("%0.2f", round(x, digits = 1)))) + 
  geom_point(stat='identity', aes(col=mpg_type), size=10)  +
  labs(y = "Average Growth Rate", x= ' Type of conviction')+
  scale_color_manual(name="Average Growth Rate of Convictions", 
                     labels = c("Increase", "Decrease"), 
                     values = c("above"="#00ba38", "below"="#f8766d")) + 
  geom_text(color="black", size=4) +
  labs(title="Regions") + 
  ylim(-9, 4) +
  coord_flip()+ theme(axis.text=element_text(size=11))

```

```{r echo=FALSE, fig.height=20, fig.width=15}
gridExtra::grid.arrange(GR_A, GR_R, nrow=2, top = textGrob("Average Growth Rate of Convictions in Auckland and Regions (1980-2014)", gp= gpar(fontsize=20, font=3)))

```
```
